\mythesistitleNOTSYNPOSIS

### 2011-12-28 src/thesis.tex
Nightly update. Trivial changes.
Cool story bro but it's FUCKING TIME TO STEP ON THE GAS.

\def \mythesissubmissiondate {30 December 2011}
\def \mythesissubmissionmonth {\MakeUppercase {December 2011}}

@@

\chapter{Preliminaries to new results}
\label{ch:prelims} 
%\section{Preliminaries} 
%\label{sec:prelims}

@@

\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH
  INCOMPLETE SEC 4.}
The problem of COP testing can be easily seen as a simple constraint
satisfaction problem involving a system of sets from a universe. Every
column of the binary matrix can be converted into a set of integers
which are the indices of the rows with $1$s in that column. When
observed in this context, if the matrix has the COP, a reordering of
its rows will result in sets that have only consecutive integers. In
other words, the sets after reordering are intervals. Indeed the
problem now becomes finding interval assignments to the given set
system \cite{nsnrs09} with a single permutation of the universe (set
of row indices) which permutes each set to its interval. The result in
\cite{nsnrs09} characterizes interval assignments to the sets which
can be obtained from a single permutation of the rows.  They show that
for each set, the cardinality of the interval assigned to it must be
same as the cardinality of the set, and the intersection cardinality
of any two sets must be same as the interesction cardinality of the
corresponding intervals.  While this is naturally a necessary
condition, \cite{nsnrs09} shows this is indeed sufficient.  Such an
interval assignment is called an Intersection Cardinality Preserving
Interval Assignment (ICPIA).  Finally, the idea of decomposing a given
0-1 matrix into prime matrices to check for COP is adopted from
\cite{wlh02} to test if an ICPIA exists for a given set system.

\tnote{TBD: have a few lines about organization of chapter}
The necessary preliminaries with definitions
etc. are presented in
Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
characterization of a feasible path labeling and finally,
Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
find the tree path labeling of a given set system from a given
$k$-subdivided tree.

\section{Introduction} 
A natural generalization of the interval
assignment problem is feasible tree path labeling problem of a set
system. The problem is defined as follows -- given a set system $\cF$
from a universe $U$ and a tree $T$, does there exist a bijection from
$U$ to the vertices of $T$ such that each set in the system maps to a
path in $T$.  We refer to this as the {\em tree path labeling problem}
for an input set system, target tree pair -- $(\cF,T)$. As a special
case if the tree $T$ is a path, the problem becomes the interval
assignment problem.  We focus on the question of generalizing the
notion of an ICPIA \cite{nsnrs09} to characterize feasible path
assignments.  We show that for a given set system $\cF$, a tree $T$,
and an assignment of paths from $T$ to the sets, there is a feasible
bijection between $U$ and $V(T)$ if and only if all intersection
cardinalities among any three sets (not necessarily distinct) is same
as the intersection cardinality of the paths assigned to them and the
input runs a filtering algorithm (described in this paper)
successfully.  This characterization is proved constructively and it
gives a natural data structure that stores all the relevant feasible
bijections between $U$ and $V(T)$.  Further, the filtering algorithm
is also an efficient algorithm to test if a tree path labeling to the
set system is feasible.  This generalizes the result in
\cite{nsnrs09}.

%\noindent 
\tnote{the following text could be in Further Study/Conclusion}It
is an interesting fact that for a matrix with the COP, the
intersection graph of the corresponding set system is an interval
graph.  A similar connection to a subclass of chordal graphs and a
superclass of interval graphs exists for the generalization of COP.
In this case, the intersection graph of the corresponding set system
must be a {\em path graph}. Chordal graphs are of great significance,
extensively studied, and have several applications.  One of the well
known and interesting properties of a chordal graphs is its connection
with intersection graphs \cite{mcg04}. For every chordal graph, there
exists a tree and a family of subtrees of this tree such that the
intersection graph of this family is isomorphic to the chordal graph
\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
called {\em clique trees} \cite{apy92} of the chordal graph. This is a
compact representation of the chordal graph. There has also been work
done on the generalization of clique trees to clique hypergraphs
\cite{km02}.  If the chordal graph can be represented as the
intersection graph of paths in a tree, then the graph is called path
graph \cite{mcg04}.  Therefore, it is clear that if there is a
bijection from $U$ to $V(T)$ such that for every set, the elements in
it map to vertices of a unique path in $T$, then the intersection
graph of the set system is indeed a path graph.  However, this is only
a necessary condition and can be checked efficiently because path
graph recognition is polynomial time
solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
system and tree, such that the intersection graph is a path graph, but
there is no bijection between $U$ and $V(T)$ such that the sets map to
paths. Path graph isomorphism is known be isomorphism-complete, see
for example \cite{kklv10}. An interesting area of research would be to
see what this result tells us about the complexity of the tree path
labeling problem (not covered in this paper).

@@

\section{ TPL with special target trees}
\tnote{give problem definition etc}

@@

\section{ TPL with no restrictions}
\tnote{ADD WALCOM PAPER TEXT}

>>>>

\def \mythesissubmissiondate {10 January 2012}
\def \mythesissubmissionmonth {\MakeUppercase {January 2012}}

@@

\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH
  INCOMPLETE SEC 4.}
As we saw earlier, tree path labeling problem of path hypergraphs is a
generalization of consecutive-ones property testing.

\remove{The problem of COP testing can be easily seen as a simple
  constraint satisfaction problem involving a system of sets from a
  universe (analogous to hypergraph). Every column of the binary
  matrix can be converted into a set of integers which are the indices
  of the rows with |un|s in that column. When observed in this
  context, if the matrix has the COP, a reordering of its rows will
  result in sets that have only consecutive integers. In other words,
  the sets after reordering are intervals. Indeed the problem now
  becomes finding interval assignments to the given set system
  cite|nsnrs09| with a single permutation of the universe (set of row
  indices) which permutes each set to its interval. The result in
  cite|nsnrs09| characterizes interval assignments to the sets which
  can be obtained from a single permutation of the rows.  They show
  that for each set, the cardinality of the interval assigned to it
  must be same as the cardinality of the set, and the intersection
  cardinality of any two sets must be same as the interesction
  cardinality of the corresponding intervals.  While this is naturally
  a necessary condition, cite|nsnrs09| shows this is indeed
  sufficient.  Such an interval assignment is called an Intersection
  Cardinality Preserving Interval Assignment (ICPIA).  Finally, the
  idea of decomposing a given 0-1 matrix into prime matrices to check
  for COP is adopted from cite|wlh02| to test if an ICPIA exists for a
  given set system.}\tnote{REPEATITION from intro chapter}

A natural generalization of the interval
assignment problem is feasible tree path labeling problem of a set
system. The problem is defined as follows -- given a set system $\cF$
from a universe $U$ and a tree $T$, does there exist a bijection from
$U$ to the vertices of $T$ such that each set in the system maps to a
path in $T$.  We refer to this as the {\em tree path labeling problem}
for an input set system, target tree pair -- $(\cF,T)$. As a special
case if the tree $T$ is a path, the problem becomes the interval
assignment problem.  We focus on the question of generalizing the
notion of an ICPIA \cite{nsnrs09} to characterize feasible path
assignments.  We show that for a given set system $\cF$, a tree $T$,
and an assignment of paths from $T$ to the sets, there is a feasible
bijection between $U$ and $V(T)$ if and only if all intersection
cardinalities among any three sets (not necessarily distinct) is same
as the intersection cardinality of the paths assigned to them and the
input runs a filtering algorithm (described in this paper)
successfully.  This characterization is proved constructively and it
gives a natural data structure that stores all the relevant feasible
bijections between $U$ and $V(T)$.  Further, the filtering algorithm
is also an efficient algorithm to test if a tree path labeling to the
set system is feasible.  This generalizes the result in
\cite{nsnrs09}.

The necessary preliminaries with definitions
etc. are presented in
Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
characterization of a feasible path labeling for any path
labeling from any target trees. Section~\ref{sec:spltargettree}
describes the special case where the target tree is of a particular
family of trees and specifically,
Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
find the tree path labeling of a given set system from a given
$k$-subdivided tree. Section~\ref{sec:norestraint} describes the
general situation where the target tree has no restrictions and the
algorithm to find a TPL, if  any, in this case.

\section{Introduction}

%\noindent 
\annote{It
is an interesting fact that for a matrix with the COP, the
intersection graph of the corresponding set system is an interval
graph.}{the following text could be in Further Study/Conclusion}  A similar connection to a subclass of chordal graphs and a
superclass of interval graphs exists for the generalization of COP.
In this case, the intersection graph of the corresponding set system
must be a {\em path graph}. Chordal graphs are of great significance,
are extensively studied, and have several applications.  One of the well
known and interesting properties of a chordal graphs is its connection
with intersection graphs \cite{mcg04}. For every chordal graph, there
exists a tree and a family of subtrees of this tree such that the
intersection graph of this family is isomorphic to the chordal graph
\cite{plr70,gav78,bp93}.  These trees when in a certain format, are
called {\em clique trees} \cite{apy92} of the chordal graph. This is a
compact representation of the chordal graph. There has also been work
done on the generalization of clique trees to clique hypergraphs
\cite{km02}.  If the chordal graph can be represented as the
intersection graph of paths in a tree, then the graph is called path
graph \cite{mcg04}.  Therefore, it is clear that if there is a
bijection from $U$ to $V(T)$ such that for every set, the elements in
it map to vertices of a unique path in $T$, then the intersection
graph of the set system is indeed a path graph.  However, this is only
a necessary condition and can be checked efficiently because path
graph recognition is polynomial time
solvable\cite{gav78,aas93}. Indeed, it is possible to construct a set
system and tree, such that the intersection graph is a path graph, but
there is no bijection between $U$ and $V(T)$ such that the sets map to
paths. Path graph isomorphism is known be isomorphism-complete, see
for example \cite{kklv10}. An interesting area of research would be to
see what this result tells us about the complexity of the tree path
labeling problem (not covered in this paper).

\section{Preliminaries to new results}
\label{ch:prelims} 
%\section{Preliminaries} 
%\label{sec:prelims}

@@

@@

\section{ TPL with special target trees}
\label{sec:spltargettree}
\tnote{give problem definition etc}

@@

\section{ TPL with no restrictions}
\label{sec:norestraint}
\tnote{ADD WALCOM PAPER TEXT}
